### 2025-06-20 index.qmd
20250619 - efficient frontier

# Automatically create a bib database for R packages
packagesForBibliography <- c(
  "petersenlab", "ffanalytics",
  "knitr", "rmarkdown", "dplyr", "ggplot2", "tidyverse", "downlit", "xml2",
  "nflreadr", "nflplotR", "nflfastR", "nflseedR", "nfl4th", "gsisdecoder", 
  "progressr", "DescTools", "pwr", "pwrss", "WebPower", "XICOR", "dagitty", 
  "ggdag", "lubridate", "ggtext", "gghighlight", "ggExtra", "grid", "patchwork", 
  "pROC",  "lme4", "lmerTest", "MuMIn", "emmeans", "pbkrtest", "sjstats", 
  "AICcmodavg", "rstan", "brms", "bbmle", "fitdistrplus", "mclust", "magrittr", 
  "viridis", "plotly", "quantmod", "fPortfolio", "xts", "zoo", "forecast", #"webshot2", 
  "stringi", "parallel", "parallelly", "doParallel", "missRanger", "ggridges", 
  "powerjoin", "caret", "gpboost", "cmdstanr", "learnr", "remotes", "corrplot", 
  "correlation", "mgcv", "rms", "car", "performance", "lavaan", "mice", 
  "miceadds", "interactions", "shiny", "PearsonDS", "robustbase", "data.table", 
  "future", "future.apply"
)

>>>>

# Automatically create a bib database for R packages
packagesForBibliography <- c(
  "petersenlab", "ffanalytics",
  "knitr", "rmarkdown", "dplyr", "ggplot2", "tidyverse", "downlit", "xml2",
  "nflreadr", "nflplotR", "nflfastR", "nflseedR", "nfl4th", "gsisdecoder", 
  "progressr", "DescTools", "pwr", "pwrss", "WebPower", "XICOR", "dagitty", 
  "ggdag", "lubridate", "ggtext", "gghighlight", "ggExtra", "grid", "patchwork", 
  "pROC",  "lme4", "lmerTest", "MuMIn", "emmeans", "pbkrtest", "sjstats", 
  "AICcmodavg", "rstan", "brms", "bbmle", "fitdistrplus", "sn", "mclust", 
  "magrittr", "viridis", "plotly", "quantmod", "fPortfolio", "NMOF", "xts", #"webshot2",
  "zoo", "forecast", "stringi", "parallel", "parallelly", "doParallel", 
  "missRanger", "ggridges", "powerjoin", "caret", "gpboost", "cmdstanr", 
  "learnr", "remotes", "corrplot", "correlation", "mgcv", "rms", "car", 
  "performance", "lavaan", "mice", "miceadds", "interactions", "shiny", 
  "PearsonDS", "robustbase", "data.table", "future", "future.apply", "SimDesign"
)

### 2025-06-20 modern-portfolio-theory.qmd
20250619 - efficient frontier

```{r}
library("quantmod")
library("fPortfolio")
library("tidyverse")
```

@@

In the example below, we use historical returns (since YEAR) as the expected future returns.
However, using historical returns as the expected future returns is risky because, as described in the common disclaimer, "Past performance does not guarantee future results."
If you select a relatively short period of historical returns, you may be selecting a period when the stock performed particularly well.
When evaluating historical returns it is preferable to evaluate long time horizons and to evaluate how the stock performed during period of both boom (i.e., "bull markets") and bust (i.e., "bear markets", such as in a recession).

@@

@@

@@

>>>>

```{r}
library("quantmod")
library("fPortfolio")
library("NMOF")
library("tidyverse")
```

### Load Data {#sec-modernPortfolioTheoryLoadData}

```{r}
#| eval: false
#| include: false

load(file = file.path(path, "/OneDrive - University of Iowa/Teaching/Courses/Fantasy Football/Data/players_projectedPoints_seasonal.RData", fsep = ""))
load(file = file.path(path, "/OneDrive - University of Iowa/Teaching/Courses/Fantasy Football/Data/player_stats_weekly.RData", fsep = ""))
load(file = file.path(path, "/OneDrive - University of Iowa/Teaching/Courses/Fantasy Football/Data/nfl_playerIDs.RData", fsep = ""))
```

```{r}
load(file = "./data/players_projectedPoints_seasonal.RData")
load(file = "./data/player_stats_weekly.RData")
load(file = "./data/nfl_playerIDs.RData")
```

@@

In the example below, we use historical returns (since 2012) as the expected future returns.
However, using historical returns as the expected future returns is risky because, as described in the common disclaimer, "Past performance does not guarantee future results."
If you select a relatively short period of historical returns, you may be selecting a period when the stock performed particularly well.
When evaluating historical returns it is preferable to evaluate long time horizons and to evaluate how the stock performed during period of both boom (i.e., "bull markets") and bust (i.e., "bear markets", such as in a recession).

@@

minYear <- format(as.Date(row.names(returns_ts)), "%Y") %>% 
  na.omit %>% 
  as.integer %>% 
  min()
```

@@

```{r}
all_proj <- dplyr::bind_rows(players_projectedPoints_seasonal)

all_proj <- all_proj %>% 
  rename(projectedPoints = raw_points)
```

```{r}
all_proj_summary <- all_proj %>% 
  group_by(id) %>% 
  summarise(
    mean = mean(projectedPoints, na.rm = TRUE),
    sd = sd(projectedPoints, na.rm = TRUE),
    var = var(projectedPoints, na.rm = TRUE)
  )

all_proj_summary <- all_proj_summary %>% 
  left_join(
    nfl_playerIDs[,c("mfl_id","name","merge_name","position","team")],
    by = c("id" = "mfl_id")
  ) %>% 
  select(name, team, position, everything()) %>% 
  arrange(-mean)
```

::: {.content-visible when-format="html:js"}

```{r}
#| label: fig-projectedPointsMeanVsSD
#| fig-cap: "Mean of Projected Fantasy Points For Each Player as a Function of the Standard Deviation of the Projected Fantasy Points Across Sources."
#| fig-alt: "Mean of Projected Fantasy Points For Each Player as a Function of the Standard Deviation of the Projected Fantasy Points Across Sources."

plot_projectedPointsMeanVsSD <- ggplot2::ggplot(
  data = all_proj_summary,
  aes(
    x = sd,
    y = mean)) +
  geom_point(
    aes(
      text = name, # add player name for mouse over tooltip
      label = position)) + # add season for mouse over tooltip
  #geom_smooth() +
  coord_cartesian(
    xlim = c(0,NA),
    ylim = c(0,NA),
    expand = FALSE) +
  labs(
    x = "Standard Deviation of Player's Projected Fantasy Points (Season) Across Sources",
    y = "Mean of Player's Projected Fantasy Points (Season) Across Sources",
    title = "Mean vs Standard Deviation of Players' Projected Fantasy Points"
  ) +
  theme_classic()

plotly::ggplotly(plot_projectedPointsMeanVsSD)
```

::: {.content-visible when-format="pdf"}

```{r}
#| label: fig-projectedPointsMeanVsSD
#| fig-cap: "Mean of Projected Fantasy Points For Each Player as a Function of the Standard Deviation of the Projected Fantasy Points Across Sources."
#| fig-alt: "Mean of Projected Fantasy Points For Each Player as a Function of the Standard Deviation of the Projected Fantasy Points Across Sources."

ggplot2::ggplot(
  data = all_proj_summary,
  aes(
    x = sd,
    y = mean)) +
  geom_point(
    aes(
      text = name, # add player name for mouse over tooltip
      label = position)) + # add season for mouse over tooltip
  #geom_smooth() +
  coord_cartesian(
    xlim = c(0,NA),
    ylim = c(0,NA),
    expand = FALSE) +
  labs(
    x = "Standard Deviation of Player's Projected Fantasy Points (Season) Across Sources",
    y = "Mean of Player's Projected Fantasy Points (Season) Across Sources",
    title = "Mean vs Standard Deviation of Players' Projected Fantasy Points"
  ) +
  theme_classic()
```

```{r}
all_proj_summary_noNA <- all_proj_summary %>% 
  filter(!is.na(id) & !is.na(mean) & !is.na(var))

all_proj_summary_noNA_removeZeroVar <- all_proj_summary_noNA %>% 
  filter(var > 0)
```

```{r}
const_cor <- function(rho, numAssets) {
  C <- array(rho, dim = c(numAssets, numAssets))
  diag(C) <- 1
  C
}

varCovMatrix <- diag(all_proj_summary_noNA_removeZeroVar$sd) %*% const_cor(0, nrow(all_proj_summary_noNA_removeZeroVar)) %*% diag(all_proj_summary_noNA_removeZeroVar$sd) # create a variance-covariance matrix assuming assets/players are uncorrelated

efficientFrontierData <- NMOF::mvFrontier(
  all_proj_summary_noNA_removeZeroVar$mean,
  varCovMatrix,
  wmin = 0,
  wmax = 1,
  n = 50)
```

```{r}
#| label: fig-efficientFrontierAcrossProjections
#| fig-cap: "Efficient Frontier of Projected Fantasy Points Across Sources."
#| fig-alt: "Efficient Frontier of Projected Fantasy Points Across Sources."

plot(
  efficientFrontierData$volatility,
  efficientFrontierData$return,
  pch = 19,
  cex = 0.5,
  type = "o",
  xlab = "Variance in Projected Fantasy Points Across Sources",
  ylab = "Projected Fantasy Points")
```

@@

```{r}
player_stats_weekly_recent <- player_stats_weekly %>% 
  filter(season == max(season))

player_stats_weekly_recent_summary <- player_stats_weekly_recent %>% 
  group_by(player_id) %>% 
  summarise(
    mean = mean(fantasyPoints, na.rm = TRUE),
    sd = sd(fantasyPoints, na.rm = TRUE),
    var = var(fantasyPoints, na.rm = TRUE)
  )

player_stats_weekly_recent_summary <- player_stats_weekly_recent_summary %>% 
  left_join(
    nfl_playerIDs[,c("gsis_id","name","merge_name","position","team")],
    by = c("player_id" = "gsis_id")
  ) %>% 
  select(name, team, position, everything()) %>% 
  arrange(-mean)
```

::: {.content-visible when-format="html:js"}

```{r}
#| label: fig-fantasyPointsMeanVsSD
#| fig-cap: "Mean of Weekly Fantasy Points For Each Player as a Function of the Week-to-Week Standard Deviation of their Fantasy Points."
#| fig-alt: "Mean of Weekly Fantasy Points For Each Player as a Function of the Week-to-Week Standard Deviation of their Fantasy Points."

plot_fantasyPointsMeanVsSD <- ggplot2::ggplot(
  data = player_stats_weekly_recent_summary,
  aes(
    x = sd,
    y = mean)) +
  geom_point(
    aes(
      text = name, # add player name for mouse over tooltip
      label = position)) + # add season for mouse over tooltip
  #geom_smooth() +
  coord_cartesian(
    xlim = c(0,NA),
    ylim = c(0,NA),
    expand = FALSE) +
  labs(
    x = "Standard Deviation of Player's Fantasy Points From Week-to-Week",
    y = "Mean of Player's Fantasy Points Across Weeks",
    title = "Mean vs Standard Deviation of Players' Fantasy Points"
  ) +
  theme_classic()

plotly::ggplotly(plot_fantasyPointsMeanVsSD)
```

::: {.content-visible when-format="pdf"}

```{r}
#| label: fig-fantasyPointsMeanVsSD
#| fig-cap: "Mean of Weekly Fantasy Points For Each Player as a Function of the Week-to-Week Standard Deviation of their Fantasy Points."
#| fig-alt: "Mean of Weekly Fantasy Points For Each Player as a Function of the Week-to-Week Standard Deviation of their Fantasy Points."

ggplot2::ggplot(
  data = player_stats_weekly_recent_summary,
  aes(
    x = sd,
    y = mean)) +
  geom_point(
    aes(
      text = name, # add player name for mouse over tooltip
      label = position)) + # add season for mouse over tooltip
  #geom_smooth() +
  coord_cartesian(
    xlim = c(0,NA),
    ylim = c(0,NA),
    expand = FALSE) +
  labs(
    x = "Standard Deviation of Player's Fantasy Points From Week-to-Week",
    y = "Mean of Player's Fantasy Points Across Weeks",
    title = "Mean vs Standard Deviation of Players' Fantasy Points"
  ) +
  theme_classic()
```

```{r}
player_stats_weekly_recent_summary_noNA <- player_stats_weekly_recent_summary %>% 
  filter(!is.na(player_id) & !is.na(mean) & !is.na(var))

player_stats_weekly_recent_summary_noNA_removeZeroVar <- player_stats_weekly_recent_summary_noNA %>% 
  filter(var > 0)
```

```{r}
varCovMatrix2 <- diag(player_stats_weekly_recent_summary_noNA_removeZeroVar$sd) %*% const_cor(0, nrow(player_stats_weekly_recent_summary_noNA_removeZeroVar)) %*% diag(player_stats_weekly_recent_summary_noNA_removeZeroVar$sd) # create a variance-covariance matrix assuming assets/players are uncorrelated

efficientFrontierData2 <- NMOF::mvFrontier(
  player_stats_weekly_recent_summary_noNA_removeZeroVar$mean,
  varCovMatrix2,
  wmin = 0,
  wmax = 1,
  n = 50)
```

```{r}
#| label: fig-efficientFrontierAcrossWeeks
#| fig-cap: "Efficient Frontier of Fantasy Points Across Weeks."
#| fig-alt: "Efficient Frontier of Fantasy Points Across Weeks."

plot(
  efficientFrontierData2$volatility,
  efficientFrontierData2$return,
  pch = 19,
  cex = 0.5,
  type = "o",
  xlab = "Variance in Fantasy Points Across Weeks",
  ylab = "Weekly Fantasy Points")
```

### 2025-06-20 multiple-regression.qmd
20250619 - efficient frontier

To examine whether residuals are normally distributed, we can examine quantile–quantile (QQ) plots and probability–probability (PP) plots.
QQ plots are particularly useful for identifying deviations from normality in the tails of the distribution; PP plots are particularly useful for identifying deviations from normality in the center of the distribution.
Researchers tend to be more concerned about the tails of the distribution, because extreme values tend to have a greater impact on inferences.
Various examples of QQ plots and deviations from normality are depicted in @fig-distributionQQplots.
If the residuals are normally distributed, they will stay close to the diagonal reference line of the QQ and PP plots.

>>>>

To examine whether residuals are normally distributed, we can examine quantile–quantile (QQ) plots and probability–probability (PP) plots.
QQ plots are particularly useful for identifying deviations from normality in the tails of the distribution; PP plots are particularly useful for identifying deviations from normality in the center of the distribution.
Researchers tend to be more concerned about the tails of the distribution, because extreme values tend to have a greater impact on inferences, so researchers tend to use QQ plots more often than PP plots.
Various examples of QQ plots and deviations from normality are depicted in @fig-distributionQQplots.
If the residuals are normally distributed, they will stay close to the diagonal reference line of the QQ and PP plots.
